Better JavaScript Runtime Understanding by Automated Function Name Extraction

### 2011-02-02 trunk/papers/issta2011/sigproc-sp.tex

To support the above argument, we analyzed the source code of five famous JavaScript projects. For every project, the total number of functions and the number of functions with name are collected (Table~\ref{functions}). The ratio of the number of functions with names to the total number of functions is less than 13 percent in all projects excepting Prototype and the average ratio is less than 5 percent. Our analysis does not include the functions defined by dynamically created scripts evaluated by {\small\texttt{eval}} function.

@@

\section{Automated Funcation Naming}
    To tackle the aforementioned function naming issues, we propose automated function naming by anlysing the source code. A function name must help the developer to
distinguish the function from other functions with different behaviours. The behaviour of a function is dependent to its source code, enclosing scopes (e.g., global and closure variables), the object bound to {\small\texttt{this}}-depending on the way the function is called-and passed arguments. The first two parameters can be almost known from the source code, while the two last parameters are usually not known until the function call at runtime. Therefore, to name a function object, it is enough to name its \textit{function body}. 
    A function body can be uniquly identified by its location in the source file. A naive proposal for the function name for a function body is a combination of the function file name and line number. Although it may work as an identifier, but it does not assist the develper to recall or understand the function behaviour. Morover, A proposed function name should be close to the name a developer would choose for the function.

Once a function object is created, one of these cases would happen: the function object is
    \begin{enumerate}
		\item assigned to a property of a new object in an object literal.
		\item assigned to a new array index in an array literal.
		\item assigned to an object property through direct access.
		\item assigned to an object property through hashmap access by a string.
		\item assigned to an object property through hashmap access by a variable name.
		\item assigned to an object property through hashmap access by a JavaScript expression.
		\item assigned to an array index.
		\item assigned to a variable.
		\item directly called.
		\item returned from a function call.
		\item passed as an argument to a function.
		\end{enumerate}

@@

	 2 &
	 & \multirow{8}{*}{\hspace{.2cm}\begin{rotate}{90}assigned to a(n)\end{rotate}}
	 & \multicolumn{2}{|m{2.8cm}|}{
	   new array index in an array literal.} 
   & foo = [..., function()\{...\}, ...] \\ 
   \cline{0-0}\cline{4-6}

   3 &
   & 
   &  
   & direct access.
   & bar.foo = function()\{...\}\\
   \cline{0-0}\cline{5-6}

   4 & \multirow{11}{*}{\hspace{.2cm}\begin{rotate}{90}The function object is\end{rotate}} 
   & 
   & 
   & \raggedright hashmap access by a string. 
   & bar["foo"] = function()\{...\} \\
   \cline{0-0}\cline{5-6}

   5 &
   & 
   & \multirow{4}{*}{\hspace{.2cm}\begin{rotate}{90}object property through\end{rotate}}
   & \raggedright hashmap access by a variable name.
   & bar[foo] = function()\{...\} \\
   \cline{0-0}\cline{5-6}

   6 &
   & 
   & 
   & \raggedright hashmap access by a JavaScript expression.
   & bar["f"+"oo"] = function()\{...\} \\
   \cline{0-0}\cline{4-6}

   7 &
   &
   & \multicolumn{2}{|l|}{
      array index.}
   & foo[0] = function()\{...\} \\
   \cline{0-0}\cline{4-6}

@@

   11 &
   & \multicolumn{3}{|m{3.5cm}|}{
     \raggedright passed as an argument to a function.}
   & foo(..., function()\{\}, ...) \\
   \hline

  \end{tabular}
  }
\end{table}

%to remove
To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debuggers and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debuggers and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debuggers and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

\subsection{Naming Functions in Call Stack}

%To choose the most appropriate name for a function which has the characteristics defined in the previous section, we consider two cases. First, assigning a global name to the function which identifies the function in the whole program. Second, naming a function in a call stack.

%A function object in JavaScript can be defined by \texttt{function \textit{identifier} (args)\{...\}} or \texttt{new Function(args, source)}.
%\subsection{A Global Name for a Function Object}

%new Function ([arg1[, arg2[, ... argN]],] functionBody)
%this[ optall.queue === false ? "each" : "queue" ](function() {})

%The object referenced by the keyword {\small\texttt{this}} is not determined by lexical scoping, but instead by the caller;
\begin{figure}[htp]
\begin{verbatim}
(a)
 foo();
(b)
 foo.bar();
(c)
 foo.apply(bar, args);
(d)
 foo.call();
\end{verbatim}
\caption{Different cases of function object creation.}
\label{fig:functionCall}
\end{figure}

\section{evaluation}
%Firebug, JQuery,

>>>>

To support the above argument, we analyzed the source code of ten famous JavaScript projects. For every project, the total number of functions and the number of functions with name are collected (Table~\ref{functions}). The ratio of the number of functions with names to the total number of functions is less than 13 percent in all projects excepting Prototype and the average ratio is less than 7 percent. Our analysis does not include the functions defined by dynamically created scripts evaluated by {\small\texttt{eval}} function. Among all functions defined in these ten projects, only 116 functions (less than 0.12 percent) are defined by the {\small\texttt{Function}} constructor.

@@

\section{Automated Funcation Naming}

    To tackle the aforementioned issues caused due to the lack of function names, we propose automated function naming by anlyzing the source code and extracting the initial name used for the function object after its creation. To illustrate the basic idea, we use the JavaScript code presented in Figure~\ref{js-code}.

    We can recognize seven function bodies in the code, none of them has a name. The first function defined in line 9 is called just after its creation and returns another function defined in line 10. The result is assigned to variable {\small\texttt{main}}. Therefore the internal function body is named {\small\texttt{main}} as its corresponding function object is assinged to a variable with this name. What about the outer function itself? We name that
 {\small\texttt{mainClosure}}, as it is used as a closure for the returned function.

@@

	 2 &
	 & \multirow{8}{*}{\hspace{.2cm}\begin{rotate}{90}assigned to a(n)\end{rotate}}
	 & \multicolumn{2}{|m{2.8cm}|}{
	   new array index in an array literal.} 
   & [..., function()\{...\}, ...] \\ 
   \cline{0-0}\cline{4-6}

   3 &
   & 
   &  
   & direct access.
   & bar\textsuperscript{*}.foo = function()\{...\}\\
   \cline{0-0}\cline{5-6}

   4 & \multirow{11}{*}{\hspace{.2cm}\begin{rotate}{90}The function object is\end{rotate}} 
   & 
   & 
   & \raggedright hashmap access by a string. 
   & bar\textsuperscript{*}["foo"] = function()\{...\} \\
   \cline{0-0}\cline{5-6}

   5 &
   & 
   & \multirow{4}{*}{\hspace{.2cm}\begin{rotate}{90}object property through\end{rotate}}
   & \raggedright hashmap access by a variable name.
   & bar\textsuperscript{*}[foo] = function()\{...\} \\
   \cline{0-0}\cline{5-6}

   6 &
   & 
   & 
   & \raggedright hashmap access by a JavaScript expression.
   & bar\textsuperscript{*}[foo\textsuperscript{*}] = function()\{...\} \\
   \cline{0-0}\cline{4-6}

   7 &
   &
   & \multicolumn{2}{|l|}{
      array index.}
   & foo\textsuperscript{*}[0] = function()\{...\} \\
   \cline{0-0}\cline{4-6}

@@

   11 &
   & \multicolumn{3}{|m{3.5cm}|}{
     \raggedright passed as an argument to a function.}
   & foo\textsuperscript{*}(..., function()\{\}, ...) \\
   \hline

  \end{tabular}
  }
 \label{function-types} 
\end{table}

 The third function (defined in line 14), is passed as an argument to another function with name {\small\texttt{op}}. We may name the function {\small\texttt{op(1)}}, as it is passed as the first argument to {\small\texttt{op}}). A better name would be the the argument name in the called function (here {\small\texttt{callback}}). However the latter requires knowning the function body of the called function which is not always possible.

    As we explained in the previous section, {\small\texttt{FOO}} and {\small\texttt{BAR}} are good name candidates for functions defined in lines 19 and 22, as these functions are assigned to variable with the same names. Similar to the first two functions which are named {\small\texttt{mainClosure}} and {\small\texttt{main}}, the functions in lines 25 and 27 can be named {\small\texttt{opClosure}} and {\small\texttt{op}} respectively.

\begin{table*}
\centering
\caption{The number of nameless functions in each category defined in table~\ref{function-types}.}
  \begin{tabular}{ | l | l | l | l | l | l | l | l | l | l | l | l |}
  \hline
     & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 \\ 
  \hline 
   Closure       & 23(0.3\%)  & 0  & 8465(94\%) & 3  & 4  & 0  & 0 & 67(0.7\%)& 16(0.2\%)& 43(0.5\%)& 365(4\%)\\ 
  \hline 
   DOJO          & 9599(61\%) & 6  & 1762(11\%) & 21 & 22 & 2  & 1 & 1143(7\%)& 471(3\%) & 169(1\%) & 2644(17\%) \\ 
  \hline 
   ExtJS         & 30221(83\%)& 0  & 1476(4\%)  & 3  & 16 & 9  & 0 & 820(2\%) & 788(2\%) & 509(1\%) & 2617(7\%) \\ 
  \hline 
   Firebug       & 2296(76\%) & 0  & 539(18\%)  & 1  & 0  & 0  & 0 & 14(0.4\%)& 7(0.2\%) & 6(0.2\%) & 149(5\%) \\ 
  \hline 
   jQuery        & 233(58\%)  & 0  & 34(9\%)    & 0  & 10 & 2  & 0 & 24(6\%)  & 9(2\%)   & 0        & 86(21\%)  \\ 
  \hline 
   MochiKit      & 1080(59\%) & 0  & 383(21\%)  & 0  & 2  & 0  & 0 & 106(6\%) & 18(1\%)  & 41(2\%)  & 181(10\%) \\ 
  \hline 
   MooTools      & 339(55\%)  & 0  & 79(13\%)   & 0  & 4  & 3  & 0 & 53(9\%)  & 21(3\%)  & 14(2\%)  & 105(17\%) \\ 
  \hline 
   Prototype     & 265(60\%)  & 0  & 28(6\%)    & 0  & 0  & 1  & 0 & 25(6\%)  & 44(10\%) & 8(2\%)   & 71(16\%) \\ 
  \hline 
   Scriptaculous & 564(64\%)  & 0  & 75(8\%)    & 0  & 2  & 2  & 0 & 27(3\%)  & 45(5\%)  & 9(1\%)   & 160(18\%) \\ 
  \hline 
   YUI           & 14154(66\%)& 7  & 1721(8\%)  & 0  & 90 & 0  & 0 & 1181(6\%)& 172(1\%) & 95(0.5\%)& 4004(19\%)\\ 
  \hline 
   All           & 58774(65\%)& 23 & 14562(16\%)& 28 & 150& 19 & 1 & 3460(4\%)& 1591(2\%)& 894(1\%) & 10382(12\%) \\ 
  \hline 
  \end{tabular}
 \label{function-types-count} 
\end{table*}

\subsection{Naming Scheme}
We can generalize the approach used for naming functions in the example by studying differnt cases that a function object can be created and used. Once a function object is created, one of the cases presented in Table~\ref{function-types} can happen. One common case of function object creation is by top script named function bodies. We did not considered this case as a separate case as it can be transformed to case 8 where a function body is assigned to a variable. Moreover, Note that functions are created by {\small\texttt{Function}} constructor can be replace by a function body and named in the same way. For every case, we will explain how a name can be assinged to the function.

Names which are specified with a star are reduced from a JavaScript expression. We discuss this reduction in ~\ref{sec:general-element-naming}. Table~\ref{function-types-count} shows the number of nameless functions in each category in the studied projects. These numbers shows how frequent and important a category of functions is in these projects. 
%they have similar pattern

\subsubsection{Case 1: in Object Literal}
 This case usually apears when developers try to group a set of functions in a new object. A common case is grouping a set of functions in the {\small\texttt{prototype}} property of the constructor. This structure resembles the class structure in traditional object-oriented languages. When a new object is created by the constructor, the new object also inherits all functions defined in the constructor's prototype. This structure is also used  when the owner object is a shared object with a set of utility functions.

 The majority of nameless functions (more than 65\%) in almost all studied projects (except Closure), are defined in object literals. For this group of functions we assign the property name used to keep a reference to the function object as the function name.

\subsubsection{Case 2: in Array Literal}
Contrary to the previous case the appearance of this case is very limited. It usually appears in initializations or when an array of functions are passed as an argument. Among all projects only two projects have instances of nameless functions defined in this way. To name the function body defined in an array literal, first we must name the array literal. To name the array literal, we consider it as a function body and we employ the same algorithm we use for naming a function body. If the array literal receives a name like {\small\texttt{foo}}, and the function is defined at index ({\small\texttt{i}}), {{\small\texttt{foo[i]}} is assigned as the function body name.

\subsubsection{Case3: setProperty by Direct Access }
This case is the second most common case in the studied projects. Here, we can see why the Closure project is different from other projects in the first case. About 94\% of nameless functions in this project are defined in this way. It seems that the Closure developers follow an internal standard for function objects creation and usage. Although most functions are assigned to objects in the initialization phase, there are still cases where functions assigned through direct access. 
This case usually happens when the functinos are assinged to object properties after its initialization. Similar to the first case the property name is assinged to the function body as its name.

\subsubsection{Case 4: setProperty by String}
In JavaScript, objects are like hashmaps and their properties can also be accessed by strings. In general an expression can be specified in the brackets after the object. The expression is evaluated at runtime and the result which should be a string is used to access the property. In the simplest form, the expression can be a string. As this form can be completely replaced by the direct access, we see few instances of this form for function object assignment. The string inside the brackets become the function name in this case.

\subsubsection{Case 5: setProperty by variable}
The usage of this case is also limited. This form usually apears when the same function body is assigned to different properties in a loop. Therefore, the variable
 name is a general name like {\small\texttt{fname}} or {\small\texttt{item}}. Morover, based on the variable value-which is not known until runtime-the owner can be a    regular object or an array. To name a function in this case we use the owner object name with the variable inside brackets like {\small\texttt{bar[foo]}}.

\subsubsection{Case 6: setProperty by Expression}
 A common case of expression in this case is a conditional expression, e.g., {\small\texttt{condtion?"prop1":"prop2"}}. To name this functions, similar to previous case we use the owner object and a sbustring of expression inside the brackets.

%jQuery.fn["inner" + name] = function() {
%dojo[dojo._defaultXhr ? "_defaultXhr" : "xhr"] = function
%scope[this.triggerEvent]

\subsubsection{Case 7: setArray}
We only observed one instance of this form in the studied projects. The function name consists of the array name with the index like {\small\texttt{foo[0]}}.

\subsubsection{Case 8: setVariable }
As the functions are immutable objects, it is very likely that a function object which is assigned to a variable, is used with the same variable name in the function scope
and its internal scopes. Although, there are cases that the variable name is temporary as the function is passed to another function or assigned to an object property, in most cases the variable name is the local function name. Therefore, for this group of functions the variable name is used as the function name.

\subsubsection{Case 9: Called}
Calling functions objects just after their creations is a common pattern in JavaScript. This pattern is usually used for building modules employing the clousure feature in JavaScript. Although the function object usage is only limited to this call, functions inside the closure function(The function is being called) has access to the properties in the closure scope in the future calls.

If the function call has no result, it means that the function performs one task (e.g., initialization of some values in the outer scope for later use). In this case, we name the function {\small\texttt{Closure}}. If the function returns a value and it is assinged to {\small\texttt{foo}}, we name the function {\small\texttt{fooClosure}}.

\subsubsection{Case 10: Returned}
Although this category only includes 2\% of nameless function, proper naming of functions in this class is important. Many constructors are build by this form and therfore the name of thes functions appear in object summaries.

If the function is returned from a function which is just called after its creation(case 9), and the result is assigned to {\small\texttt{foo}}, we name the function {\small\texttt{foo}}. Otherwise, if the name of outer function is {\small\texttt{foo}}, we name the function {\small\texttt{fooResult}}.

% this is also important as many constructor/classes are made in this way

\subsubsection{Case 11: as an Argument}
Numbers in table~\ref{function-types-count} show that creating and passing functions as arguments is very common in JavaScript. If the calling function name is 
{\small\texttt{foo}} and the function is passed as the {\small\texttt{ith}} argument, we name the function {\small\texttt{foo(i)}}. One of the common cases of passing
functions as arguments is to register function in a global shared object. The function is usually registered with a name which is also passed as string argument. Therefore, to imporove our naming approach, we include the first string argument (if there is any) in the name. For example if the string argument is {\small\texttt{myFunction}} the assigned name will be {\small\texttt{foo(i)(myFunction)}}.

% to register a function with a name
% to wrap the function
% as a callback

\subsection{Reducing a JavaScript Expression to a Name}
\label{sec:general-element-naming}

\begin{table}
\centering
\caption{JavaScript Expression Reduction to Name.}
\scalebox{0.9}{
  \begin{tabular}{ | l | l | l |}
  \hline
   Description & Total & Named \\ 
  \hline 
   Primitive Value & value & value.toString() \\ 
  \hline 
   Variable & foo & foo \\ 
  \hline 
   GetProp & bar.foo & Name(bar).foo \\ 
  \hline 
   GetElem & bar[foo] & Name(bar)+[Name(foo)] \\ 
  \hline 
   Add & bar op foo & Name(bar)+op+Name(foo) \\ 
  \hline 
   Condition & cond?bar:foo & Name(bar)+":"+Name(foo) \\ 
  \hline 
  \end{tabular}
  }
 \label{expression-reduction} 
\end{table}

\subsection{Similar Function Names in a File}
It is conceivable, that function bodies are named with the same names in a file. It can happen in different situations. One of the common cases is the two different function bodies may assing to a property based on a condition (e.g., the browser). For these cases we add a dollar sign and an index to the similar function names. For example, if there are two function bodies with names {\small\texttt{myFunction}}, the functions are named as {\small\texttt{myFunction\$0}} and {\small\texttt{myFunction\$1}}.

%examples: toString functions in different classes, myFunction = cond? function(){..} : function(){..} 
% more than one return

\subsection{Package Name}

\subsection{Functions created by {\large \texttt{eval()}}}}
The function bodies in {\small\texttt{eval}} statements can be named similar to other objects as the source code is avaialble. However, eval scripts have no file name and therefore, it causes issues for assigning appropriate packages name to these functions.

\section{Improved by Runtime Data}
\subsection{Naming Functions in Call Stack}

\begin{figure}[htp]
\begin{verbatim}
(a)
 foo();
(b)
 foo.bar();
(c)
 foo.apply(bar, args);
(d)
 foo.call();
\end{verbatim}
\caption{Different cases of function object creation.}
\label{fig:functionCall}
\end{figure}

\section{evaluation}

\subsection{Developers' vs. Automated Names}

\subsection{Compared to Firebug {\large \texttt{guessFunctionName()}}}
Firebug uses a utility function called {\small \texttt{guessFunctionName()}} to name nameless functions.
This algorithm used in the function searches lines around the line the function definition line for specific patterns by regular expressions. The
  first string matches one of the patterns becomes the name of the function. We compared the results of our algorithm to Firebug's in table~\ref{firebug}.

The first column shows, how many functions get similar names by both algorithm. In average, for 60\% of functions both algorithms produce the same names.
Firebug is not able to name about 15\% of functions. In 35\% of functions the results are different. In most cases, the proposed name by Firebug is the name of another function and is wrong.

\begin{table}
\centering
\caption{Number of similar, and different function names in comparison to Firebug {\small \texttt{guessFunctionName()}}.}
\scalebox{0.9}{
  \begin{tabular}{ | l | l | l | l | l |}
  \hline
   Project & Similar & Different & UnNamed \\ 
  \hline 
   Closure & 8524(1\%) & 9195 & 516(2\%) \\ 
  \hline 
   DoJo & 12180(1\%) & 9195 & 5780(2\%) \\ 
  \hline 
   ExtJS & 100(1\%) & 9195 & 208(2\%) \\ 
  \hline 
   Firebug & 2901(1\%) & 9195 & 494(0\%) \\ 
  \hline 
   jQuery & 283(1\%) & 9195 & 117(0\%) \\ 
  \hline 
   MochiKit &  1568(1\%) & 9195 & 234(2\%) \\ 
  \hline 
   MooTools & 440(1\%) & 9195 & 128(2\%) \\ 
  \hline 
   Prototype & 292(1\%) & 9195 & 326(0\%) \\ 
  \hline 
   Scriptaculous &  100(1\%) & 9195 & 397(2\%) \\ 
  \hline 
   YUI &  11248(1\%) & 9195 & 3702(2\%) \\ 
  \hline 
   All & 100(1\%) & 9195 & 208(2\%) \\ 
  \hline 
  \end{tabular}
  }
 \label{firebug} 
\end{table}

\subsection{Manual Check}
%Firebug, JQuery,
